# Causality

## Prerequistes

This chapter only uses the tidyverse package
```{r}
library("tidyverse")
```

## Racial Discrimination in the Labor Market

The code in the book uses `table` and `addmargins` to construct the table.
However, this can be done easily with `dplyr` using grouping and summarizing.

```{r message=FALSE}
data("resume", package = "qss")
```

In addition to the functions shown in the text,
```{r}
dim(resume)
summary(resume)
head(resume)
```
we can also use `glimpse` to get a quick understanding of the variables in the data frame,
```{r}
glimpse(resume)
```

For each combination of `race` and `call` let's count the observations:
```{r}
race_call_tab <-
  resume %>%
  group_by(race, call) %>%
  count()
race_call_tab
```

If we want to calculate callback rates by race:
```{r}
race_call_rate <-
  race_call_tab %>%
  group_by(race) %>%
  mutate(call_rate =  n / sum(n)) %>%
  filter(call == 1) %>%
  select(race, call_rate)
race_call_rate
```

If we want the overall callback rate, we can calculate it from the original 
data,
```{r}
resume %>%
  summarise(call_back = mean(call))
```

## Subsetting Data in R


### Subsetting

The **dplyr** function `filter` is a much improved version of `subset`.

To select black individuals in the data:
```{r}
resumeB <-
  resume %>%
  filter(race == "black")
dim(resumeB)
head(resumeB)
```

And to calculate the callback rate
```{r}
resumeB %>%
  summarise(call_rate = mean(call))
```

To keep call and first name variables and those with black-sounding first names.
```{r}
resumeBf <-
  resume %>%
  filter(race == "black", sex == "female") %>%
  select(call, firstname)
head(resumeBf)
```

Now we can calculate the gender gap by group. 

One way to do this is to calculate the call back rates for both sexes of 
black sounding names,
```{r}
resumeB <- resume %>%
  filter(race == "black") %>%
  group_by(sex) %>%
  summarise(black_rate = mean(call))
```
and white-sounding names
```{r}
resumeW <- resume %>%
  filter(race == "white") %>%
  group_by(sex) %>%
  summarise(white_rate = mean(call))
resumeW
```
Then, merge `resumeB` and `resumeW` on `sex` and calculate the difference for both sexes.
```{r}
inner_join(resumeB, resumeW, by = "sex") %>%
  mutate(race_gap = white_rate - black_rate)
  
```

This seems to be a little more code, but we didn't duplicate as much as in QSS, and this would easily scale to more than two categories.

A way to do this using the `spread` and gather functions from **tidy** are,
First, calculate the 
```{r}
resume_race_sex <-
  resume %>%
  group_by(race, sex) %>%
  summarise(call = mean(call))
head(resume_race_sex)
```
Now, use `spread()` to make each value of `race` a new column:
```{r}
library("tidyr")
resume_sex <-
  resume_race_sex %>%
  ungroup() %>%
  spread(race, call)
resume_sex
```
Now we can calculate the race wage differences by sex as before,
```{r}
resume_sex %>%
  mutate(call_diff = white - black)
```

This could be combined into a single chain with only six lines of code:
```{r}
resume %>%
  group_by(race, sex) %>%
  summarise(call = mean(call)) %>%
  ungroup() %>%
  spread(race, call) %>%
  mutate(call_diff = white - black)
```



### Simple conditional statements

See the **dlpyr** functions `if_else`, `recode` and `case_when`.
The function `if_else` is like `ifelse` but corrects for some weird behavior that `ifelse` has in certain cases.

```{r}
resume %>% 
  mutate(BlackFemale = if_else(race == "black" & sex == "female", 1, 0)) %>%
  group_by(BlackFemale, race, sex) %>%
  count()
```


### Factor Variables

See R4DS Chapter 15 "Factors" and the package **forcats**

The code in this section works, but can be simplified by using the function
`case_when` which works in exactly these cases.
```{r}
resume %>%
  mutate(type = as.factor(case_when(
    .$race == "black" & .$sex == "female" ~ "BlackFemale",
    .$race == "black" & .$sex == "male" ~ "BlackMale",
    .$race == "white" & .$sex == "female" ~ "WhiteFemale",
    .$race == "white" & .$sex == "male" ~ "WhiteMale",
    TRUE ~ as.character(NA)
  )))
```

Since the logic of this is so simple, we can create this variable by 
using `str_c` to combine the vectors of `sex` and `race`, after using `str_to_title` to capitalize them first.
```{r}
library(stringr)
resume <-
  resume %>%
  mutate(type = str_c(str_to_title(race), str_to_title(sex)))
```

Some of the reasons given for using factors in this chapter are not as important given the functionality in modern **tidyverse** packages.
For example, there is no reason to use `tapply`, as that can use `group_by` and `summarise`,
```{r}
resume %>%
  group_by(type) %>%
  summarise(call = mean(call))
```

What's nice about this approach is that we wouldn't have needed to create the factor variable first,
```{r}
resume %>%
  group_by(race, sex) %>%
  summarise(call = mean(call))
```


We can use that same approach to calculate the mean of first names, and use
`arrange` to sort in ascending order.
```{r}
resume %>%
  group_by(firstname) %>%
  summarise(call = mean(call)) %>%
  arrange(call)
```

## Causal Affects and the Counterfactual
 
Load the data using the **readr** function `r RDoc("readr::read_csv")`
```{r message=FALSE}
data("social", package = "qss")
summary(social)
```

Use a grouped summarize instead of `tapply`,
```{r}
gotv_by_group <-
  social %>%
  group_by(messages) %>%
  summarize(turnout = mean(primary2006))
gotv_by_group
```

Get the turnout for the control group
```{r}
gotv_control <-
  (filter(gotv_by_group, messages == "Control"))[["turnout"]]
```

Subtract the control group turnout from all groups

```{r}
gotv_by_group %>%
  mutate(diff_control = turnout - gotv_control)
```

We could have also done this in one step like,
```{r}
gotv_by_group %>%
  mutate(control = mean(turnout[messages == "Control"]),
         control_diff = turnout - control)
```

We can compare the differences of variables across the groups easily using a grouped summarize
```{r}
gotv_by_group %>%
  mutate(control = mean(turnout[messages == "Control"]),
         control_diff = turnout - control)
```


**Pro-tip** The `summarise_at` functions allows you summarize one-or-more columns with one-or-more functions.
In addition to `age`, 2004 turnout, and household size, we'll also compare proportion female,
```{r}
social %>%
  group_by(messages) %>%
  mutate(age = 2006 - yearofbirth,
         female = (sex == "female")) %>%
  select(-age, -sex) %>%
  summarise_all(mean)
```

## Observational Studies

Load the `minwage` dataset from its URL using `readr::read_csv`:
```{r message=FALSE}
minwage_url <- "https://raw.githubusercontent.com/kosukeimai/qss/master/CAUSALITY/minwage.csv"
minwage <- read_csv(minwage_url)
glimpse(minwage)
summary(minwage)
```

First, calculate the proportion of restaurants by state whose hourly wages were less than the minimum wage in NJ, \$5.05, for `wageBefore` and `wageAfter`:

Since the NJ minimum wage was \$5.05, we'll define a variable with that value.
Even if you use them only once or twice, it is a good idea to put values like this in variables. 
It makes your code closer to self-documenting.n
```{r}
NJ_MINWAGE <- 5.05
```
Later, it will be easier to understand `wageAfter < NJ_MINWAGE` without any comments than it would be to understand `wageAfter < 5.05`. 
In the latter case you'd have to remember that the new NJ minimum wage was 5.05 and that's why you were using that value.
This is an example of a [magic number](https://en.wikipedia.org/wiki/Magic_number_(programming)#Unnamed_numerical_constants): try to avoid them.

Note that location has multiple values: PA and four regions of NJ.
So we'll add a state variable to the data.
```{r}
minwage %>%
  count(location)
```

We can extract the state from the final two characters of the location variable using the **stringr** function `str_sub` (R4DS Ch 14: Strings):
```{r}
library(stringr)
minwage <-
  mutate(minwage, state = str_sub(location, -2L))
```
Alternatively, since everything is either PA or NJ
```{r eval=FALSE}
minwage <-
  mutate(minwage, state = if_else(location == "PA", "PA", "NJ"))
```

Let's confirm that the restaurants followed the law:
```{r}
minwage %>%
  group_by(state) %>%
  summarise(prop_after = mean(wageAfter < NJ_MINWAGE),
            prop_Before = mean(wageBefore < NJ_MINWAGE))
```

Create a variable for the proportion of full-time employees in NJ and PA 
```{r}
minwage <-
  minwage %>%
  mutate(totalAfter = fullAfter + partAfter,
        fullPropAfter = fullAfter / totalAfter)
```

Now calculate the average for each state:
```{r}
full_prop_by_state <-
  minwage %>%
  group_by(state) %>%
  summarise(fullPropAfter = mean(fullPropAfter))
full_prop_by_state
```

We could compute the difference by  
```{r}
(filter(full_prop_by_state, state == "NJ")[["fullPropAfter"]] - 
  filter(full_prop_by_state, state == "PA")[["fullPropAfter"]])
```
or using **tidyr** functions `spread` (R4DS Ch 11: Tidy Data):
```{r}
spread(full_prop_by_state, state, fullPropAfter) %>%
  mutate(diff = NJ - PA)
```



### Confounding Bias

We can calculate the proportion of fast-food restaurants in each chain in each state:
```{r}
chains_by_state <-
  minwage %>%
  group_by(state) %>%
  count(chain) %>%
  mutate(prop = n / sum(n)) 
```

We can easily compare these using a simple dot-plot:
```{r}
ggplot(chains_by_state, aes(x = chain, y = prop, colour = state)) +
  geom_point() + 
  coord_flip()
```

In the QSS text, only Burger King restaurants are compared. 
However, **dplyr** makes this easy.
All we have to do is change the `group_by` statement we used last time,
and add chain to it:

```{r}
full_prop_by_state_chain <-
  minwage %>%
  group_by(state, chain) %>%
  summarise(fullPropAfter = mean(fullPropAfter))
full_prop_by_state_chain
```

We can plot and compare the proportions easily in this format.
In general, ordering categorical variables alphabetically is useless, so we'll order the chains by the average of the NJ and PA `fullPropAfter`, using `forcats::fct_reorder`:
```{r}
ggplot(full_prop_by_state_chain,
       aes(x = forcats::fct_reorder(chain, fullPropAfter),
           y = fullPropAfter, 
           colour = state)) +
  geom_point() +
  coord_flip() +
  labs(x = "chains")
```

To calculate the differences, we need to get the data frame 

1. The join method.

   1. Create New Jersey and Pennsylvania data sets with `chain` and prop full employed columns.
   2. Merge the two data sets on `chain`.
   
```{r}
chains_nj <- full_prop_by_state_chain %>%
  ungroup() %>%
  filter(state == "NJ") %>%
  select(-state) %>%
  rename(NJ = fullPropAfter)
chains_pa <- full_prop_by_state_chain %>%
  ungroup() %>%
  filter(state == "PA") %>%
  select(-state) %>%
  rename(PA = fullPropAfter)

full_prop_state_chain_diff <- 
  full_join(chains_nj, chains_pa, by = "chain") %>%
  mutate(diff = NJ - PA)
full_prop_state_chain_diff
```

Q: In the code above why did I remove the `state` variable and rename the `fullPropAfter` variable before merging? What happens if I didn't?

2. The spread/gather method. We can also use the `spread` and `gather` functions from **tidyr**. In this example it is much more compact code.

```{r}
full_prop_by_state_chain %>%
  spread(state, fullPropAfter) %>%
  mutate(diff = NJ - PA)
```

### Before and After and Difference-in-Difference Designs


To compute the estimates in the before and after design first create a variable for the difference before and after the law passed.
```{r}
minwage <- 
  minwage %>%
  mutate(totalBefore = fullBefore + partBefore,
         fullPropBefore = fullBefore / totalBefore)
```

The before-and-after analysis is the difference between the full-time employment before and after the minimum wage law passed looking only at NJ:
```{r}
filter(minwage, state == "NJ") %>%
  summarise(diff = mean(fullPropAfter) - mean(fullPropBefore))
```

The difference-in-differences design uses the difference in the before-and-after differences for each state.
```{r}
diff_by_state <-
  minwage %>%
  group_by(state) %>%
  summarise(diff = mean(fullPropAfter) - mean(fullPropBefore))

filter(diff_by_state, state == "NJ")[["diff"]] -
  filter(diff_by_state, state == "PA")[["diff"]]
```

Let's create a single dataset with the mean values of each state before and after to visually look at each of these designs:
```{r}
full_prop_by_state <-
  minwage %>%
  group_by(state) %>%
  summarise_at(vars(fullPropAfter, fullPropBefore), mean) %>%
  gather(period, fullProp, -state) %>%
  mutate(period = recode(period, fullPropAfter = 1, fullPropBefore = 0))
full_prop_by_state
```

```{r}
ggplot(full_prop_by_state, aes(x = period, y = fullProp, colour = state)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = c(0, 1), labels = c("Before", "After"))
```



## Descriptive Statistics for a Single Variable


To calculate the summary for the variables `wageBefore` and `wageAfter`:
```{r}
minwage %>%
  filter(state == "NJ") %>%
  select(wageBefore, wageAfter) %>%
  summary()
  
```

We calculate the IQR for each state's wages after the passage of the law using the same grouped summarize as we used before:
```{r}
minwage %>%
  group_by(state) %>%
  summarise(wageAfter = IQR(wageAfter),
            wageBefore = IQR(wageBefore))
```

Calculate the variance and standard deviation of `wageAfter` and `wageBefore` for each state:

```{r}
minwage %>%
  group_by(state) %>%
  summarise(wageAfter_sd = sd(wageAfter),
               wageAfter_var = var(wageAfter),
               wageBefore_sd = sd(wageBefore),
               wageBefore_var = var(wageBefore))

```
or, more compactly, using `summarise_at`:
```{r}
minwage %>%
  group_by(state) %>%
  summarise_at(vars(wageAfter, wageBefore), funs(sd, var))
```

