---
title: "QSS Tidyverse Code"
author: "Jeffrey B. Arnold"
date: "`r Sys.Date()`"
github-repo: jrnold/qss-tidy
site: "bookdown::bookdown_site"
documentclass: book
---

# Preface

This is tidyverse R code to supplement the book, [Quantitative Social Science: An Introduction](http://press.princeton.edu/titles/11025.html), by Kosuke Imai, to
be published by Princeton University Press in March 2017.

The R code included with the text of QSS and the supplementary materials relies mostly on base R functions. 
This translates the code examples provided with QSS to tidyverse R code. 
[Tidyverse](https://github.com/tidyverse/tidyverse) refers to a set of packages (**ggplot2**, **dplyr**, **tidyr**, **readr**, **purrr**, **tibble**,  and a few others) that share common data representations, especially the use of data frames for return values. The book [R for Data Science](http://r4ds.had.co.nz/) by Hadley Wickham and Garrett Grolemond is an introduction. 

To install the R packages used in this work run the following code:
```{r eval=FALSE}
# install.packages("devtools")
install_github("jrnold/qss-tidy")
```
It install's the **qsstidy** package which contains no code or data, but will install the needed dependencies.

I wrote this code while teaching course that employed both texts in order to make the excellent examples and statistical material in QSS more compatible with the modern data science R approach in R4DS.

## Colonphon

The source of the book is available [here](https://github.com/jrnold/qsstidy) and was built with versions of packages below:

```{r colophon, echo = FALSE}
devtools::session_info()
```
