---
knit: "bookdown::render_book"
title: "Quantatitive Social Science: The R Tidyverse Code"
description: >
  This supplement for Kosuke Imai's "Quantitative Social Science: An Introduction"
  contains Tidyverse R code versions of the code in the original text.
author: "Jeffrey B. Arnold"
twitter: jrnld
github-repo: "jrnold/qss-tidy"
url: 'http\://jrnold.github.io/qss-tidy'
date: "`r Sys.Date()`"
site: "bookdown::bookdown_site"
documentclass: book
---

# Preface {-}

This is tidyverse R code to supplement the book, [Quantitative Social Science: An Introduction](http://press.princeton.edu/titles/11025.html), by Kosuke Imai.

The R code included with the text of *QSS* and the supplementary materials relies mostly on base R functions.
This translates the code examples provided with *QSS* to tidyverse R code.
[Tidyverse](https://github.com/tidyverse/tidyverse) refers to a set of packages (**ggplot2**, **dplyr**, **tidyr**, **readr**, **purrr**, **tibble**,  and a few others) that share common data representations, especially the use of data frames for return values.

This is not a complete introduction to R and the tidyverse.
I suggest pairing it with [R for Data Science](http://r4ds.had.co.nz/) by Hadley Wickham and Garrett Grolemond.

These materials are supplement to replace the existing *QSS* code with the tidvyerse dialect of R.
Thus it does not replicate the substantive material, and not meant to be used independently of the *QSS* text.
However, the provided code is not merely a translation of the *QSS* code.
It often uses the topics in *QSS* to delve deeper into data science, data visualization, and computational topics.

I wrote this code while teaching course that employed both texts in order to make the excellent examples and statistical material in *QSS* more compatible with the modern data science using R approach in *R4DS*.

## Colophon {-}

To install the R packages used in this work run the following code, installs the **qsstidy** package which contains no code or data, but will install the needed dependencies.
```{r eval=FALSE}
install.packages("devtools")
install_github("jrnold/qss-tidy")
```

Additionally, the `r pkg("gganimate")` package requires installing [ffmpeg](https://ffmpeg.org/) with libvpx support.

The source of the book is available [here](https://github.com/jrnold/qsstidy) and was built with versions of packages below:

```{r colophon, echo = FALSE}
devtools::session_info()
```
