# Introduction

## Prerequisites

```{r}
library("tidyverse")
```
We also load the **haven** package to load Stata `dta` files,
```{r}
library("haven")
```
and the **rio** package to load multiple types of files.
```{r}
library("rio")
```
We also create a helper function for loading in the data.
We will also, use the function `qss_data_url`:
```{r echo=FALSE}
insert_fun("qss_data_url")
```
```{r qss_data_url-source,eval=FALSE}
```

This function returns the URL to a data set.
```{r}
qss_data_url("intro", "Kenya.csv")
```
Since URLs can be used for paths to files, this can be used to easily load data directly from the site:
```{r message=FALSE}
read_csv(qss_data_url("intro", "Kenya.csv"))
```



## Introduction to R

### Data Files

Don't use `setwd()` within scripts.
It is much better to organize your code in projects.

When reading from csv files use `readr::read_csv` instead of the base R function `read.csv`. 
It is slightly faster, and returns a `tibble` instead of a data frame.
See `r R4DS` [Ch 11: Data Import](http://r4ds.had.co.nz/tibbles.html#introduction-4)
for more discussion.

We also can load it directly from a URL.
```{r}
UNpop <- read_csv(qss_data_url("INTRO", "UNpop.csv"))
class(UNpop)
UNpop
```


Datasets can also be distributed with R packages. 
These are often smaller datasets used in examples and package tutorials.

The function `data` with called without any arguments will list all the datasets distributed with installed packages.
```{r eval=FALSE}
data()
```
The **qss** library distributes the data sets used in the *QSS* textbook.
```{r}
data("UNpop", package = "qss")
```

The single bracket, `[`, is useful to select rows and columns in simple cases,
but the **dplyr** functions `slice()` to select rows by number, `filter` to select by certain criteria, or `select()` to select columns.

```{r}
UNpop[c(1, 2, 3), ]
```
is equivalent to 
```{r}
UNpop %>% slice(1:3)
```

```{r}
UNpop[, "world.pop"]
```
is almost equivalent to 
```{r}
select(UNpop, world.pop)
```
However, note that `select()` **always** returns a tibble, and never a vector, 
even if only one column is selected.
Also, note that since `world.pop` is a tibble, using `[` also returns tibbles 
rather than a vector if it is only one column.

```{r}
UNpop[1:3, "year"]
```
is equivalent to 
```{r}
UNpop %>%
  slice(1:3) %>%
  select(year)
```
For this example using these functions and `%>%` to chain them together may seem verbose, but later we can produce more complicated transformations of the data by chaining together simple functions.

```{r}
UNpop$world.pop[seq(from = 1, to = nrow(UNpop), by = 2)]
```
can be rewritten as
```{r}
UNpop %>%
  slice(seq(1, n(), by = 2)) %>%
  select(world.pop) 
```

The function `n()` when used in a **dplyr** function returns the number of rows
in the data frame (or the number of rows in the group if used with `group_by`).



### Saving Objects

**Do not save** the work space using `save.image`.
This is an extremely bad idea for reproducibility.

See the `r R4DS` chapter [Workflow Projects](http://r4ds.had.co.nz/workflow-projects.html). 
You should uncheck the options in RStudio to avoid saving and restoring from `.RData` files. 
This will help ensure that your R code runs the way you think it does, instead of depending on some long forgotten code that is only saved in the workspace image. 

Everything important should be in a script. Anything saved or loaded from file
should be done explicitly.

As with reading CSV files, use the **readr** package functions. 
In this case, `write_csv` writes a csv file
```{r eval=FALSE}
write_csv(UNpop, "UNpop.csv")
```


### Packages

To read and write Stata and SPSS 

```{r}
read_dta(qss_data_url("INTRO", "UNpop.dta"))
```

There is also the equivalent `write_dta` function to create Stata datasets.
```{r, eval=FALSE}
write_dta(UNpop, "UNpop.dta")
```

Also see the [rio](https://cran.r-project.org/package=rio) package which makes loading data even easier with smart defaults.

You can use the `import` function to load many types of files:
```{r}
import(qss_data_url("INTRO", "UNpop.csv"))
import(qss_data_url("INTRO", "UNpop.RData"))
import(qss_data_url("INTRO", "UNpop.dta"))
```

R also includes the **foreign** package, which contains functions for reading and writing  **haven**. 
One reason to do so is that it is better maintained. 
For  R function `read.dta` does not read files created by the most recent versions of Stata (13+).


### Style Guide


In addition to `r pkg("lintr")` the R package `r pkg("formatR")` has methods to clean up your code.

