---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Introduction

## Prerequisites {-}

In the prerequisites section of each chapter, we'll load any packages
needed for the chapter, and possibly define some functions or load data.


```{r}
library("tidyverse")
```
We also load the **readr** package to load `csv` files,

```{r}
library("readr")
```
the **haven** package to load Stata `dta` files,

```{r}
library("haven")
```
and the **rio** package to load multiple types of files.

```{r}
library("rio")
```



## Overview of the Book

<!-- define all sections so that numbering matches that of the book -->


## How to use the Book


## Introduction to R

These notes do not aim to completely teach R and the tidyverse. 
However, there are many other resources for that.

[R for Data Science](http://r4ds.had.co.nz/) is a comprehensive introduction to R 
using the tidyverse.

[Data Camp](https://www.datacamp.com/home) has interactive courses. In particular,
I recommend starting with the following two courses.

- [Introduction to R](https://www.datacamp.com/courses/free-introduction-to-r)
- [Introduction to the Tidyverse](https://www.datacamp.com/courses/introduction-to-the-tidyverse)


### Arithmetic Operations

See *QSS* text


### Objects

See *QSS* text. Also see [R4DS: Workflow basics](http://r4ds.had.co.nz/workflow-basics.html).


### Vectors

See *QSS* text. Also see [R4DS: Vectors](http://r4ds.had.co.nz/vectors.html). In 
*R for Data Science* vectors are introduced much later, after data frames.


### Functions

See *QSS* text. Also see [R4DS: Functions](http://r4ds.had.co.nz/functions.html).


### Data Files


Rather than using `setwd()` in scripts, data analysis should be organized in
projects. Read the introduction on RStudio projects in  [R4DS](http://r4ds.had.co.nz/workflow-projects.html).[^setwd]


[^setwd]: For more on using projects read [Project-oriented workflow](https://www.tidyverse.org/articles/2017/12/workflow-vs-script/).


Datasets used in R are accessed in two ways.

- Datasets can be distributed with R packages. These are often smaller datasets used in examples and tutorials in packages. These are loaded with the `data()` function. For example you can load UN data on demographic statistics from the **qss** library, which distributes the data sets used in the *QSS* textbook. (The function `data()` called without any arguments will list all the datasets distributed with installed packages.)

```{r}
data("UNpop", package = "qss")
```
- Datasets can be loaded from external files including both stored R objects (`.RData`, `.rda`) and other formats (`.csv`, `.dta`, `.sav`). To read a [csv](https://en.wikipedia.org/wiki/Comma-separated_values) file into R use the `read_csv` function from the **readr** library, part of the tidyverse.

```{r}
UNpop_URL <- "https://raw.githubusercontent.com/kosukeimai/qss/master/INTRO/UNpop.csv"
UNpop <- read_csv(UNpop_URL)
```

We use the readr function`read_csv()` instead of the base R function `read.csv()` used in the *QSS* text. It is slightly faster, and returns a `tibble` instead of a data frame. Check this by calling `class()` on the new object.

```{R}
class(UNpop)
UNpop
```
See [R for Data Science](http://r4ds.had.co.nz/) [Ch 11: Data Import](http://r4ds.had.co.nz/tibbles.html#introduction-4)
for more discussion.

Note that in the previous code we loaded the file directly from a URL, but we could also work with local files on your computer, e.g.
```{r eval=FALSE}
UNpop <- read_csv("INTRO/UNpop.csv")
```

See [R for Data Science](http://r4ds.had.co.nz/) [Ch 10: Tibbles](http://r4ds.had.co.nz/tibbles.html) for a deeper discussion of data frames.

The single bracket, `[`, is useful to select rows and columns in simple cases.

```{r}
UNpop[c(1, 2, 3), ]
```

There are **dplyr** functions to select rows by number, to select rows by certain criteria, or to select columns.

To select rows 1--3, use `slice()`.

```{r}
slice(UNpop, 1:3)
```

Base R allows you to choose the column `world.pop` column from the `UNpop` data frame:

```{r}
UNpop[, "world.pop"]
UNpop$world.pop
UNpop[["world.pop"]]
select(UNpop, world.pop)
```
Unlike `[`, the `[[` and `$` operators can only select a single column and return a vector.[^extract1]
The `dplyr` function `select()` **always** returns a tibble (data frame), and never a vector, even if only one column is selected.

[^extract1]: See the discussion in [R for Data Science](http://r4ds.had.co.nz/tibbles.html#tibbles-vs.data.frame) on how `tibble` objects differ from base `data.frame` objects in how `[` is handled.
  

Select rows 1--3 of the `year` column:
```{r}
UNpop[1:3, "year"]
```
or,
```{r}
select(slice(UNpop, 1:3), year)
```

The same series of functions can be performed using the pipe operator, `%>%`.

```{r}
UNpop %>%
  slice(1:3) %>%
  select(year)
```
This example may seem verbose, but later we can produce more complicated transformations of the data by chaining together simple functions.

Select every other row from `UNpop`:

```{r}
UNpop$world.pop[seq(from = 1, to = nrow(UNpop), by = 2)]
```
or

```{r}
UNpop %>%
  slice(seq(1, n(), by = 2)) %>%
  select(world.pop)
```
or
```{r}
UNpop %>%
  filter(row_number() %% 2 == 1)
```

The function `n()` when used in a **dplyr** function returns the number of rows in the data frame (or the number of rows in the group if used with `group_by()`).
The function  `row_number()` returns the row number of an observation. 
The `%%` operator returns the modulus, i.e. division remainder.


### Saving Objects

It is not recommended that you save the entire R workspace using `save.image` due to the negative and unexpected impacts it can have on reproducibility.See the [R for Data Science](http://r4ds.had.co.nz/) chapter [Workflow Projects](http://r4ds.had.co.nz/workflow-projects.html).

You should uncheck the options in RStudio to avoid saving and restoring from `.RData` files (go to `Tools > Global Options > General`).
This will help ensure that your R code runs the way you think it does, instead of depending on some long forgotten code that is only saved in the workspace image.
Everything important should be in a script. Anything saved or loaded from file
should be done explicitly.

Your motto should be that the **source is real**, not the objects created by it.

> The source code is real. The objects are realizations of the source code. Source for EVERY user modified object is placed in a particular directory or directories, for later editing and retrieval. – from the [ESS manual](https://ess.r-project.org/Manual/ess.html#Philosophies-for-using-ESS_0028S_0029)

This means that while you should not save the entire workplace it is perfectly fine practice to run a script and save or load R objects to files, using  or .

As with reading CSV files, use the [readr](http://readr.tidyverse.org/) package functions.
In this case, `write_csv()` writes a csv file and takes at least two objects: the data that you want to write to a csv and the name that you want to give the file.

```{r}
write_csv(UNpop, "UNpop.csv")
```


### Programming and Learning Tips

Use the [haven](http://haven.tidyverse.org/) package to read and write Stata (`.dta`) and SPSS (`.sav`) files.
Stata and SPSS are two other statistical programs commonly used in social science.
Even if you don't ever use them, you'll almost certainly encounter data stored in their native formats.

```{r}
UNpop_dta_url <- "https://github.com/kosukeimai/qss/raw/master/INTRO/UNpop.dta"
UNpop <- read_dta(UNpop_dta_url)
UNpop
```

There is also the equivalent `write_dta` function to create Stata datasets.

```{r}
write_dta(UNpop, "UNpop.dta")
```

While Stata and SPSS data sets are quite similar to data frames, they differ slightly in definitions of acceptable data types of columns and what metadata they store with the data.
 +Be careful when reading and writing from these formats to ensure that information is not lost.

Also see the [rio](https://cran.r-project.org/package=rio) package which makes loading data even easier with smart defaults.

You can use the `import` function to load many types of files:

```{r}
import("https://github.com/kosukeimai/qss/raw/master/INTRO/UNpop.csv")

import("https://github.com/kosukeimai/qss/raw/master/INTRO/UNpop.RData")

import("https://github.com/kosukeimai/qss/raw/master/INTRO/UNpop.dta")
```

R also includes the **foreign** package, which contains functions for reading and writing files using **haven**.
One reason to use these packages is that they are better maintained.
For example, the R function `read.dta` does not read files created by the most recent versions of Stata (13+), whereas **haven** does.


### Style Guide

Following a consistent coding style is important for your code to be readable by you and others.
The preferred style is the [tidyverse style guide](http://style.tidyverse.org/), which
differs slightly from [Google's R style guide](http://style.tidyverse.org/).

- The [lintr](https://cran.r-project.org/package=lintr) package will check files for style errors.
- The [styler](https://cran.r-project.org/package=styler) package provides functions for automatically formatting
    R code according to style guides.
- In RStudio, go to the `Tools > Global Options > Code > Diagnostics` pane and check the
    box to activate style warnings. On this pane, there are other options that can be 
    set in order to increase or decrease the amount of warnings while writing R code in RStudio.
