# Measurement

## Prerequisites

```{r message=FALSE}
library("tidyverse")
library("forcats")
library("broom")
```

## Measuring Civilian Victimization during Wartime

```{r message=FALSE}
afghan <- read_csv(qss_data_url("measurement", "afghan.csv"))
```

Summarize the variables of interest
```{r}
afghan %>%
  select(age, educ.years, employed, income) %>%
  summary()
```

With `income`, `r RDoc("readr::read_csv")` never converts strings to factors by default.
To get a summary of the different levels, either convert it to a factor (R4DS Ch 15), or use `count()`
```{r}
afghan %>%
  count(income)
```

Count the number a proportion of respondents who answer that they were harmed by the ISF (`violent.exp.ISAF`) and (`violent.exp.taliban`) respectively,
```{r}
afghan %>%
  group_by(violent.exp.ISAF, violent.exp.taliban) %>%
  count() %>%
  ungroup() %>%
  mutate(prop = n / sum(n))
```
We need to use `ungroup()` in order to ensure that `sum(n)` sums over the entire
dataset as opposed to only within categories of `violent.exp.ISAF`.


Unlike `prop.table`, the code above does not drop missing values.
We can drop those values by adding a `filter` verb and using `!is.na()` to test
for missing values in those variables:
```{r}
afghan %>%
  filter(!is.na(violent.exp.ISAF), !is.na(violent.exp.taliban)) %>%
  group_by(violent.exp.ISAF, violent.exp.taliban) %>%
  count() %>%
  ungroup() %>%
  mutate(prop = n / sum(n))
```

### Handling Missing Data in R

We already observed the issues with `NA` values in calculating the proportion
answering the "experienced violence" questions.

You can filter rows with specific variables having missing values using `filter`
as shown above.

However, `na.omit` works with tibbles just like any other data frame.
```{r}
na.omit(afghan)
```

## Visualizing the Univariate Distribution 

### Barplot

```{r}
library(forcats)
afghan <-
  afghan %>%
  mutate(violent.exp.ISAF.fct = 
           fct_explicit_na(fct_recode(factor(violent.exp.ISAF),
                                      Harm = "1", "No Harm" = "0"),
                           "No response"))
ggplot(afghan, aes(x = violent.exp.ISAF.fct, y = ..prop.., group = 1)) +
  geom_bar() +
  xlab("Response category") +
  ylab("Proportion of respondents") +
  ggtitle("Civilian Victimization by the ISAF")
```

```{r}
afghan <-
  afghan %>%
  mutate(violent.exp.taliban.fct = 
           fct_explicit_na(fct_recode(factor(violent.exp.taliban),
                                      Harm = "1", "No Harm" = "0"),
                           "No response"))
ggplot(afghan, aes(x = violent.exp.ISAF.fct, y = ..prop.., group = 1)) +
  geom_bar() +
  xlab("Response category") +
  ylab("Proportion of respondents") +
  ggtitle("Civilian Victimization by the Taliban")
```


This plot could improved by plotting the two values simultaneously to be able to better compare them.

This will require creating a data frame that has the following columns: perpetrator (`ISAF`, `Taliban`), response (`No Harm`, `Harm`, `No response`). 


```{r}
library("tidyr")
violent_exp <-
  afghan %>%
  select(violent.exp.ISAF, violent.exp.taliban) %>%
  gather(perpetrator, response) %>%
  mutate(perpetrator = str_replace(perpetrator, "violent\\.exp\\.", ""),
         perpetrator = str_replace(perpetrator, "taliban", "Taliban"),
         response = fct_recode(factor(response), c("Harm" = "1", "No Harm" = "0")),
         response = fct_explicit_na(response, "No response"),
         response = fct_relevel(response, c("No response", "No Harm"))) %>%
  count(perpetrator, response) %>%
  mutate(prop = n / sum(n))
         
ggplot(violent_exp, aes(x = prop, y = response, color = perpetrator)) +
  geom_point() +
  scale_color_manual(values = c(ISAF = "green", Taliban = "black"))

```

Black was chosen for the Taliban, and Green for ISAF because it is the color of their respective [flags](https://en.wikipedia.org/wiki/International_Security_Assistance_Force).


### Boxplot

```{r}
ggplot(afghan, aes(x = 1, y = age)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "Age", x = "") +
  ggtitle("Distribution of Age")
```


```{r}
ggplot(afghan, aes(y = educ.years, x = province)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = "Province", y = "Years of education") +
  ggtitle("Education by Province")

```

Helmand and Uruzgan have much lower levels of education than the other
provinces, and also report higher levels of violence.
```{r}
afghan %>%
  group_by(province) %>%
  summarise(educ.years = mean(educ.years, na.rm = TRUE),
            violent.exp.taliban =
              mean(violent.exp.taliban, na.rm = TRUE),
            violent.exp.ISAF =
              mean(violent.exp.ISAF, na.rm = TRUE)) %>%
  arrange(educ.years)
```

### Printing and saving graphics

Use the function `ggsave()` to save **ggplot2** graphics. 

Also, R Markdown files have their own means of creating and saving plots
created by code-chunks.


## Survey Sampling

### The Role of Randomization

## load village data

```{r message=FALSE}
afghan.village <- read_csv(qss_data_url("measurement", "afghan-village.csv"))
```

Box-plots of altitude
```{r}
ggplot(afghan.village, aes(x = factor(village.surveyed,
                                      labels = c("sampled", "non-sampled")),
                           y = altitude)) +
  geom_boxplot() +
  labs(y = "Altitude (meter)", x = "") +
  coord_flip()

```

Box plots log-population values of sampled and non-sampled
```{r}
ggplot(afghan.village, aes(x = factor(village.surveyed,
                                      labels = c("sampled", "non-sampled")),
                           y = log(population))) +
  geom_boxplot() +
  labs(y = "log(population)", x = "") +
  coord_flip()
```

You can also compare these distributions by plotting their densities:
```{r}
ggplot(afghan.village, aes(colour = factor(village.surveyed,
                                      labels = c("sampled", "non-sampled")),
                           x = log(population))) +
  geom_density() +
  geom_rug() +
  labs(x = "log(population)", colour = "")
```
The function `r ggdoc("geom_rug")`, creates a rug plot, which puts small lines on the axis to represent the value of each observation.
It can be combined with a scatter or density plot to add extra detail.

### Non-response and other sources of bias

Calculate the rates of non-response by province to `violent.exp.ISAF` and
`violent.exp.taliban`:
```{r}
afghan %>%
  group_by(province) %>%
  summarise(ISAF = mean(is.na(violent.exp.ISAF)),
            taliban = mean(is.na(violent.exp.taliban))) %>%
  arrange(-ISAF)
```


Calculate the proportion who support the ISAF using the difference in means
between the ISAF and control groups:
```{r}
(mean(filter(afghan, list.group == "ISAF")$list.response) -
  mean(filter(afghan, list.group == "control")$list.response))
```


To calculate the table responses to the list experiment in the control, ISAF,
and Taliban groups>
```{r}
afghan %>%
  group_by(list.response, list.group) %>%
  count() %T>%
  glimpse() %>%
  spread(list.group, n, fill = 0)
```

## Measuring Political Polarization

```{r message=FALSE}
congress <- read_csv(qss_data_url("measurement", "congress.csv"))
```

```{r}
glimpse(congress)
```

```{r}
q <- 
  congress %>%
  filter(congress %in% c(80, 112), 
         party %in% c("Democrat", "Republican")) %>%
  ggplot(aes(x = dwnom1, y = dwnom2, colour = party)) +
  geom_point() +
  facet_wrap(~ congress) +
  coord_fixed() +
  scale_y_continuous("racial liberalism/conservatism",
                     limits = c(-1.5, 1.5)) +
  scale_x_continuous("economic liberalism/conservatism",
                     limits = c(-1.5, 1.5))
q
```

However, since there are colors associated with Democrats (blue) and Republicans (blue), we should use them rather than the defaults.
There's some evidence that using semantically-resonant colors can help decoding data visualizations ([Lin, et al. 2013](http://vis.stanford.edu/files/2013-SemanticColor-EuroVis.pdf)).

Since I'll reuse the scale several times, I'll save it in a variable.
```{r}
scale_colour_parties <-
  scale_colour_manual(values = c(Democrat = "blue",
                                 Republican = "red",
                                 Other = "green"))
q + scale_colour_parties
```


```{r fig.width = 7, fig.height = 7}
congress %>%
  ggplot(aes(x = dwnom1, y = dwnom2, colour = party)) +
  geom_point() +
  facet_wrap(~ congress) +
  coord_fixed() +
  scale_y_continuous("racial liberalism/conservatism",
                     limits = c(-1.5, 1.5)) +
  scale_x_continuous("economic liberalism/conservatism",
                     limits = c(-1.5, 1.5)) +
  scale_colour_parties
```

```{r}
congress %>%
  group_by(congress, party) %>%
  summarise(dwnom1 = mean(dwnom1)) %>%
  filter(party %in% c("Democrat", "Republican")) %>%
  ggplot(aes(x = congress, y = dwnom1, 
             colour = fct_reorder2(party, congress, dwnom1))) +
  geom_line() +
  scale_colour_parties +
  labs(y = "DW-NOMINATE score (1st Dimension)", x = "Congress",
       colour = "Party")
```



### Correlation

Let's plot the Gini coefficient
```{r message=FALSE}
USGini <- read_csv(qss_data_url("measurement", "USGini.csv"))
```

```{r}
ggplot(USGini, aes(x = year, y = gini)) +
  geom_point() +
  geom_line() +
  labs(x = "Year", y = "Gini coefficient") +
  ggtitle("Income Inequality")
```

To calculate a measure of party polarization take the code used in the plot of Republican and Democratic party median ideal points and adapt it to calculate the difference in the party medians:
```{r}
party_polarization <- 
  congress %>%
  group_by(congress, party) %>%
  summarise(dwnom1 = mean(dwnom1)) %>%
  filter(party %in% c("Democrat", "Republican")) %>%
  spread(party, dwnom1) %>%
  mutate(polarization = Republican - Democrat)
party_polarization
```

```{r}
ggplot(party_polarization, aes(x = congress, y = polarization)) +
  geom_point() +
  geom_line() +
  ggtitle("Political Polarization") +
  labs(x = "Year", y = "Republican median − Democratic median")
```



### Quantile-Quantile Plot

```{r}
congress %>%
  filter(congress == 112, party %in% c("Republican", "Democrat")) %>%
  ggplot(aes(x = dwnom2, y = ..density..)) +
  geom_histogram() +
  facet_grid(party ~ .) + 
  labs(x = "racial liberalism/conservatism dimension")
```

*ggplot2* includes a `stat_qq` which can be used to create qq-plots but it is more suited to comparing a sample distribution with a theoretical distribution, usually the normal one.
However, we can calculate one by hand, which may give more insight into exactly what the qq-plot is doing.

```{r}
party_qtiles <- tibble(
  probs = seq(0, 1, by = 0.01),
  Democrat = quantile(filter(congress, congress == 112, party == "Democrat")$dwnom2,
         probs = probs),
  Republican = quantile(filter(congress, congress == 112, party == "Republican")$dwnom2,
         probs = probs)
)
party_qtiles
```

The plot looks different than the one in the text since the x- and y-scales are in the original values instead of z-scores (see the next section).
```{r}
party_qtiles %>%
  ggplot(aes(x = Democrat, y = Republican)) + 
  geom_point() +
  geom_abline() +
  coord_fixed()
```

## Clustering


### Matrices

While matrices are great for numerical computations, such as when you are 
implementing algorithms, generally keeping data in data frames is more convenient for data wrangling.

### Lists 

See R4DS [Chapter 20: Vectors](http://r4ds.had.co.nz/vectors.html),  [Chapter 21: Iteration](http://r4ds.had.co.nz/iteration.html) and the **purrr** package for more powerful methods of computing on lists.

### k-means algorithms

**TODO** A good visualization of the k-means algorithm and a simple, naive implementation in R.

Calculate the clusters by the 80th and 112th congresses,
```{r}
k80two.out <- 
  kmeans(select(filter(congress, congress == 80),
                       dwnom1, dwnom2),
              centers = 2, nstart = 5)
```

Add the cluster ids to data sets
```{r}
congress80 <- 
  congress %>%
  filter(congress == 80) %>%
  mutate(cluster2 = factor(k80two.out$cluster))
```

We will also create a data sets with the cluster centroids.
These are in the `centers` element of the cluster object.
```{r}
k80two.out$centers
```
To make it easier to use with **ggplot2**, we need to convert this to a data frame.
The `tidy` function from the **broom** package:
```{r}
k80two.clusters <- tidy(k80two.out)
k80two.clusters
```


Plot the ideal points and clusters
```{r}
ggplot() +
  geom_point(data = congress80,
             aes(x = dwnom1, y = dwnom2, colour = cluster2)) +
  geom_point(data = k80two.clusters, mapping = aes(x = x1, y = x2))

```

We can also plot,
```{r}
congress80 %>%
  group_by(party, cluster2) %>%
  count()
```

And now we can repeat these steps for the 112th congress:
```{r}
k112two.out <-
  kmeans(select(filter(congress, congress == 112),
                dwnom1, dwnom2),
         centers = 2, nstart = 5)

congress112 <-
  filter(congress, congress == 112) %>%
  mutate(cluster2 = factor(k112two.out$cluster))

k112two.clusters <- tidy(k112two.out)

ggplot() +
  geom_point(data = congress112,
             mapping = aes(x = dwnom1, y = dwnom2, colour = cluster2)) +
  geom_point(data = k112two.clusters,
             mapping = aes(x = x1, y = x2))

```

```{r}
congress112 %>%
  group_by(party, cluster2) %>%
  count()
```

Now repeat the same with four clusters on the 80th congress:
```{r}
k80four.out <-
  kmeans(select(filter(congress, congress == 80),
                dwnom1, dwnom2),
         centers = 4, nstart = 5)

congress80 <-
  filter(congress, congress == 80) %>%
  mutate(cluster2 = factor(k80four.out$cluster))

k80four.clusters <- tidy(k80four.out)

ggplot() +
  geom_point(data = congress80,
             mapping = aes(x = dwnom1, y = dwnom2, colour = cluster2)) +
  geom_point(data = k80four.clusters,
             mapping = aes(x = x1, y = x2), size = 3)

```

and on the 112th congress:
```{r}
k112four.out <-
  kmeans(select(filter(congress, congress == 112),
                dwnom1, dwnom2),
         centers = 4, nstart = 5)

congress112 <-
  filter(congress, congress == 112) %>%
  mutate(cluster2 = factor(k112four.out$cluster))

k112four.clusters <- tidy(k112four.out)

ggplot() +
  geom_point(data = congress112,
             mapping = aes(x = dwnom1, y = dwnom2, colour = cluster2)) +
  geom_point(data = k112four.clusters,
             mapping = aes(x = x1, y = x2), size = 3)

```
