---
output: html_document
editor_options:
  chunk_output_type: console
---
# Prediction


## Prerequisites {-}

```{r message = FALSE}
library("tidyverse")
library("lubridate")
library("stringr")
library("forcats")
```
The packages `r pkg_link("modelr")` and `r pkg_link("broom")` are used to wrangle the results of linear regressions,
```{r message=FALSE}
library("broom")
library("modelr")
```

## Predicting Election Outcomes

### Loops in R


RStudio provides many features to help debugging, which will be useful in
for loops and function: see  [this](https://support.rstudio.com/hc/en-us/articles/205612627-Debugging-with-RStudio) article for an example.


```{r}
values <- c(2, 3, 6)
n <- length(values)
results <- rep(NA, n)
for (i in 1:n) {
  results[i] <- values[i] * 2
  cat(values[i], "times 2 is equal to", results[i], "\n")
}
```

Note that the above code uses the for loop for pedagogical purposes only, this could have simply been written
```{r}
results <- values * 2
results
```
In general, avoid using for loops when there is a *vectorized* function.

But sticking with the for loop, there are several things that could be improved.

Avoid using the idiom `1:n` in for loops. 
To see why, look what happens when values are empty:
```{r}
values <- c()
n <- length(values)
results <- rep(NA, n)
for (i in 1:n) {
  cat("i = ", i, "\n")
  results[i] <- values[i] * 2
  cat(values[i], "times 2 is equal to", results[i], "\n")
}
```
Instead of not running a loop, as you would expect, it runs two loops, where `i = 1`, then `i = 0`.
This edge case occurs more than you may think, especially if you are writing functions
where you don't know the length of the vector is *ex ante*.

The way to avoid this is to use either `rdoc("base", "seq_len")` or `rdoc("base", "seq_along")`, which will handle 0-length vectors correctly.
```{r}
values <- c()
n <- length(values)
results <- rep(NA, n)
for (i in seq_along(values)) {
  results[i] <- values[i] * 2
}
print(results)
```
or 
```{r}
values <- c()
n <- length(values)
results <- rep(NA, n)
for (i in seq_len(n)) {
  results[i] <- values[i] * 2
}
print(results)
```

Also, note that the the result is `logical(0)`.
That's because the `NA` missing value has class [logical](http://r4ds.had.co.nz/vectors.html#missing-values-4), and thus `rep(NA, ...)` returns a logical vector.
It is better style to initialize the vector with the same data type that you will be using,
```{r}
results <- rep(NA_real_, length(values))
results
class(results)
```


Often loops can be rewritten to use a map function. 
Read the `r R4DS` chapter [Iteration](http://r4ds.had.co.nz/data-visualisation.html) before proceeding.

To do so, we first write a function that will be applied to each element of the vector.
When converting from a `for` loop to a function, this is usually simply the body of the `for` loop, though you 
may need to add arguments for any variables defined outside the body of the for loop.
In this case,
```{r}
mult_by_two <- function(x) {
  x * 2
}
```
We can now test that this function works on different values:
```{r}
mult_by_two(0)
mult_by_two(2.5)
mult_by_two(-3)
```

At this point, we could replace the body of the `for` loop with this function:
```{r}
values <- c(2, 4, 6)
n <- length(values)
results <- rep(NA, n)
for (i in seq_len(n)) {
  results[i] <- mult_by_two(values[i])
}
print(results)
```
This can be useful if the body of a for loop is many lines long.

However, this loop is still unwieldy code. We have to remember to define an empty vector `results` that is the same size as `values` to hold the results, and then correctly loop over all the values. 
We already saw how these steps have possibilities for errors. 
Functionals like `map`, apply a function to each element of a vector. 
```{r}
results <- map(values, mult_by_two)
results
```

The values of each element are correct, but `map` returns a list vector, not a numeric vector like we may have been expecting. 
If we want a numeric vector, use `map_dbl`,
```{r}
results <- map_dbl(values, mult_by_two)
results
```

Also, instead of explicitly defining a function, like `mult_by_two`, we could have instead used an *anonymous function* with the functional.
An anonymous function is a function that is not assigned to a name.
```{r}
results <- map_dbl(values, function(x) x * 2)
results
```
The various `r pkg("purrr")` functions also will interpret formulas as functions where `.x` and `.y` are interpreted as (up to) two arguments.
```{r}
results <- map_dbl(values, ~ .x * 2)
results
```
This is for parsimony and convenience; in the background, these functions are creating anonymous functions from the given formula.

*QSS* discusses several debugging strategies. The functional approach lends itself to easier debugging because the function can be tested with input values independently of the loop.


### General Conditional Statements in R

See the *R for Data Science* section [Conditional Execution](http://r4ds.had.co.nz/functions.html#conditional-execution) for a more complete discussion of conditional execution.

If you are using conditional statements to assign values for data frame,
see the **dplyr** functions `r rdoc("dplyr", "if_else")`, `r rdoc("dplyr", "recode")`, and `r rdoc("dplyr", "case_when")`

The following code which uses a for loop, 
```{r}
values <- 1:5
n <- length(values)
results <- rep(NA_real_, n)
for (i in seq_len(n)) {
  x <- values[i]
  r <- x %% 2
  if (r == 0) {
    cat(x, "is even and I will perform addition", x, " + ", x, "\n")
    results[i] <- x + x
  } else {
    cat(x, "is even and I will perform multiplication", x, " * ", x, "\n")
    results[i] <- x * x
  }
}
results
```
could be rewritten to use `if_else`,
```{r}
if_else(values %% 2 == 0, values + values, values * values)
```
or using the `map_dbl` functional with a named function,
```{r}
myfunc <- function(x) {
  if (x %% 2 == 0) {
    x + x
  } else {
    x * x
  }
}
map_dbl(values, myfunc)
```
or `map_dbl` with an anonymous function,
```{r}
map_dbl(values, function(x) {
  if (x %% 2 == 0) {
    x + x
  } else {
    x * x
  }
})
```



### Poll Predictions

Load the election polls by state for the 2008 US Presidential election,
```{r polls08,message=FALSE}
data("polls08", package = "qss")
glimpse(polls08)
```
and the election results,
```{r pres08,message=FALSE}
data("pres08", package = "qss")
glimpse(pres08)
```

Compute Obama's margin in polls and final election
```{r polls08_margin}
polls08 <-
  polls08 %>% mutate(margin = Obama - McCain)
pres08 <-
  pres08 %>% mutate(margin = Obama - McCain)
```

To work with dates, the R package `r pkg_link("lubridate")` makes wrangling them
much easier.
See the `r R4DS` chapter [Dates and Times](http://r4ds.had.co.nz/dates-and-times.html).

The function `r rdoc("lubridate", "ymd")` will convert character strings like `year-month-day` and more
into dates, as long as the order is (year, month, day). See `r rdoc("lubridate", "dmy")`, `r rdoc("lubridate", "mdy")`, and others for other ways to convert strings to dates.
```{r}
x <- ymd("2008-11-04")
y <- ymd("2008/9/1")
x - y
```

However, note that in `polls08`, the date `middate` is *already* a `date` object,
```{r}
class(polls08$middate)
```
The function `r rdoc("readr", "read_csv")` by default will check character vectors to see if they have patterns that appear to be dates, and if so, will
parse those columns as dates.

We'll create a variable for election day
```{r}
ELECTION_DAY <- ymd("2008-11-04")
```
and add a new column to `poll08` with the days to the election
```{r}
polls08 <- mutate(polls08, ELECTION_DAY - middate)
```

Although the code in the chapter uses a `for` loop, there is no reason to do so.
We can accomplish the same task by merging the election results data to the polling data by `state`.

```{r}
polls_w_results <- left_join(polls08,
                            select(pres08, state, elec_margin = margin),
                            by = "state") %>%
  mutate(error = elec_margin - margin)
glimpse(polls_w_results)
```

To get the last poll in each state, arrange and filter on `middate`
```{r}
last_polls <-
  polls_w_results %>%
  arrange(state, desc(middate)) %>%
  group_by(state) %>%
  slice(1)
last_polls
```


**Challenge:** Instead of using the last poll, use the average of polls in the last week? Last month? How do the margins on the polls change over the election period?

To simplify things for later, let's define a function `rmse` which calculates the root mean squared error, as defined in the book.
See the `r R4DS` chapter [Functions](http://r4ds.had.co.nz/functions.html) for more on writing functions.

```{r}
rmse <- function(actual, pred) {
  sqrt(mean( (actual - pred) ^ 2))
}
```
Now we can use `rmse()` to calculate the RMSE for all the final polls:
```{r}
rmse(last_polls$margin, last_polls$elec_margin)
```
Or since we already have a variable `error`,
```{r}
sqrt(mean(last_polls$error ^ 2))
```
The mean prediction error is
```{r}
mean(last_polls$error)
```


This is slightly different than what is in the book due to the difference in the poll used as the final poll; many states have many polls on the last day.

I'll choose bin widths of 1%, since that is fairly interpretable:
```{r}
ggplot(last_polls, aes(x = error)) +
  geom_histogram(binwidth = 1, boundary = 0)
```

The text uses bin widths of 5%:
```{r}
ggplot(last_polls, aes(x = error)) +
  geom_histogram(binwidth = 5, boundary = 0)
```

**Challenge:** What other ways could you visualize the results? How would you show all states? What about plotting the absolute or squared errors instead of the errors?

**Challenge:** What happens to prediction error if you average polls?
Consider averaging back over time?
What happens if you take the averages of the state poll average and average of **all** polls - does that improve prediction?

To create a scatter plots using the state abbreviations instead of points use
`r ggdoc("geom_text")` instead of `r ggdoc("geom_point")`.
```{r}
ggplot(last_polls, aes(x = margin, y = elec_margin, label = state)) +
  geom_abline(color = "white", size = 2) +
  geom_hline(yintercept = 0, color = "gray", size = 2) +
  geom_vline(xintercept = 0, color = "gray", size = 2) +
  geom_text() +
  coord_fixed() +
  labs(x = "Poll Results", y = "Actual Election Results")
```

We can create a confusion matrix as follows.
Create a new column `classification` which shows how the poll's classification was related to the actual election outcome ("true positive", "false positive", "true negative", "false negative").
If there were two outcomes, then we would use the `r RDoc("dplyr::if_then")` function.
But with more than two outcomes, it is easier to use the `r pkg("dplyr")` function `r RDoc("dplyr::case_when")`.
```{r}
last_polls <-
  last_polls %>%
  ungroup() %>%
  mutate(classification =
           case_when(
             (.$margin > 0 & .$elec_margin > 0) ~ "true positive",
             (.$margin > 0 & .$elec_margin < 0) ~ "false positive",
             (.$margin < 0 & .$elec_margin < 0) ~ "true negative",
             (.$margin < 0 & .$elec_margin > 0) ~ "false negative"
           ))
```
You need to use `.` to refer to the data frame when using `case_when()` within `mutate()`.
Also, we needed to first use `r RDoc("ungroup")` in order to remove the grouping variable so `mutate()` will work.

Now simply count the number of polls in each category of `classification`:
```{r}
last_polls %>%
  group_by(classification) %>%
  count()
```

Which states were incorrectly predicted by the polls, and what was their margins?
```{r}
last_polls %>%
  filter(classification %in% c("false positive", "false negative")) %>%
  select(state, margin, elec_margin, classification) %>%
  arrange(desc(elec_margin))
```

What was the difference in the poll prediction of electoral votes and actual electoral votes?
We hadn't included the variable `EV` when we first merged, but that's no problem, we'll just merge again in order to grab that variable:
```{r}
last_polls %>%
  left_join(select(pres08, state, EV), by = "state") %>%
  summarise(EV_pred = sum( (margin > 0) * EV),
            EV_actual = sum( (elec_margin > 0) * EV))
```


```{r message=FALSE}
data("pollsUS08", package = "qss")
```
```{r}
pollsUS08 <- mutate(pollsUS08, DaysToElection = ELECTION_DAY - middate)
```

We'll produce the seven-day averages slightly differently than the method used in the text.
For all dates in the data, we'll calculate the moving average.
The code presented in *QSS* uses a for loop similar to the following:
```{r}
all_dates <- seq(min(polls08$middate), ELECTION_DAY, by = "days")

# Number of poll days to use
POLL_DAYS <- 7

pop_vote_avg <- vector(length(all_dates), mode = "list")
for (i in seq_along(all_dates)) {
  date <- all_dates[i]
  # summarise the seven day
  week_data <-
     filter(polls08,
            as.integer(middate - date) <= 0,
            as.integer(middate - date) > - POLL_DAYS) %>%
     summarise(Obama = mean(Obama, na.rm = TRUE),
               McCain = mean(McCain, na.rm = TRUE))
  # add date for the observation
  week_data$date <- date
  pop_vote_avg[[i]] <- week_data
}

pop_vote_avg <- bind_rows(pop_vote_avg)
```

Write a function which takes a `date`, and calculates the `days` (set the default to 7 days) moving average using the dataset `.data`:
```{r}
poll_ma <- function(date, .data, days = 7) {
  filter(.data,
        as.integer(middate - date) <= 0,
        as.integer(middate - date) > - !!days) %>%
  summarise(Obama = mean(Obama, na.rm = TRUE),
           McCain = mean(McCain, na.rm = TRUE)) %>%
  mutate(date = !!date)
}
```
The code above uses `!!`. 
This tells `filter` that `days` refers to a variable `days` in the calling environment,
and not a column named `days` in the data frame. 
In this case, there wouldn't be any ambiguities since there is not a column named `days`, but in general
there can be ambiguities in the dplyr functions as to whether the names refer to columns in the data frame
or variables in the environment calling the function.
Read [Programming with dplyr](http://dplyr.tidyverse.org/articles/programming.html) for an in-depth 
discussion of this.

This returns a one row data frame with the moving average for McCain and Obama on  Nov 1, 2008.
```{r}
poll_ma(as.Date("2008-11-01"), polls08)
```
Since we made `days` an argument to the function we could easily change the code to calculate
other moving averages,
```{r}
poll_ma(as.Date("2008-11-01"), polls08, days = 3)
```
Now use a functional to execute that function with all dates for which we want moving averages.
The function `poll_ma` returns a data frame, and our ideal output is a data frame that 
stacks those data frames row-wise. 
So we will use the `map_df` function,
```{r echo=TRUE,results='hide'}
map_df(all_dates, poll_ma, polls08)
```
Note that the other arguments for `poll_ma` are placed after the name of the function as additional arguments to `map_df`.

It is easier to plot this if the data are tidy, with `Obama` and `McCain` as categories of a column `candidate`.
```{r}
pop_vote_avg_tidy <-
  pop_vote_avg %>%
  gather(candidate, share, -date, na.rm = TRUE)
head(pop_vote_avg_tidy)
```

```{r}
ggplot(pop_vote_avg_tidy, aes(x = date, y = share,
                              colour = fct_reorder2(candidate, date, share))) +
  geom_point() +
  geom_line() +
  scale_colour_manual("Candidate",
                      values = c(Obama = "blue", McCain = "red"))
```


**Challenge** read `r R4DS` chapter [Iteration](http://r4ds.had.co.nz/iteration.html#the-map-functions) and use the function `r rdoc("purrr", "map_df")` to create the object `poll_vote_avg` as above instead of a for loop.

The 7-day average is similar to the simple method used by [Real Clear Politics](http://www.realclearpolitics.com/epolls/2016/president/us/general_election_trump_vs_clinton-5491.html).
The RCP average is simply the average of all polls in their data for the last seven days.
Sites like [538](https://fivethirtyeight.com) and the [Huffpost Pollster](http://elections.huffingtonpost.com/pollster), on the other hand, also use what amounts to averaging polls, but using more sophisticated statistical methods to assign different weights to different polls.

**Challenge** Why do we need to use different polls for the popular vote data? Why not simply average all the state polls?
What would you have to do?
Would the overall popular vote be useful in predicting state-level polling, or vice-versa? How would you use them?


## Linear Regression

### Facial Appearance and Election Outcomes

Load the `face` dataset:
```{r face,message=FALSE}
data("face", package = "qss")
```
Add Democrat and Republican vote shares, and the difference in shares:
```{r face_shares}
face <- mutate(face,
                d.share = d.votes / (d.votes + r.votes),
                r.share = r.votes / (d.votes + r.votes),
                diff.share = d.share - r.share)
```

Plot facial competence vs. vote share:
```{r}
ggplot(face, aes(x = d.comp, y = diff.share, colour = w.party)) +
  geom_ref_line(h = 0) +
  geom_point() +
  scale_colour_manual("Winning\nParty",
                      values = c(D = "blue", R = "red")) +
  labs(x = "Competence scores for Democrats",
       y = "Democratic margin in vote share")
```

### Correlation and Scatter Plots

```{r}
cor(face$d.comp, face$diff.share)
```


### Least Squares

Run the linear regression
```{r}
fit <- lm(diff.share ~ d.comp, data = face)
fit
```

There are many functions to get data out of the `lm` model.

In addition to these, the `r pkg("broom")` package provides three functions: `glance`, `tidy`, and `augment` that always return data frames.

The function `r rdoc("broom", "glance.lm", text="glance")` returns a one-row data-frame summary of the model,
```{r}
glance(fit)
```
The function `r rdoc("broom", "tidy.lm", text="tidy")` returns a data frame in which each row is a coefficient,
```{r}
tidy(fit)
```
The function `r rdoc("broom", "augment.lm", text="augment")` returns the original data with fitted values, residuals, and other observation level stats from the model appended to it.
```{r}
augment(fit) %>% head()
```


We can plot the results of the bivariate linear regression as follows:
```{r}
ggplot() +
  geom_point(data = face, mapping = aes(x = d.comp, y = diff.share)) +
  geom_ref_line(v = mean(face$d.comp)) +
  geom_ref_line(h = mean(face$diff.share)) +
  geom_abline(slope = coef(fit)["d.comp"],
              intercept = coef(fit)["(Intercept)"],
              colour = "red") +
  annotate("text", x = 0.9, y = mean(face$diff.share) + 0.05,
           label = "Mean of Y", color = "blue", vjust = 0) +
  annotate("text", y = -0.9, x = mean(face$d.comp), label = "Mean of X",
           color = "blue", hjust = 0) +
  scale_y_continuous("Democratic margin in vote shares",
                     breaks = seq(-1, 1, by = 0.5), limits = c(-1, 1)) +
  scale_x_continuous("Democratic margin in vote shares",
                     breaks = seq(0, 1, by = 0.2), limits = c(0, 1)) +
  ggtitle("Facial compotence and vote share")
```

A more general way to plot the predictions of the model against the data
is to use the methods described in [Ch 23.3.3](http://r4ds.had.co.nz/model-basics.html#visualising-models) of R4DS.
Create an evenly spaced grid of values of `d.comp`, and add predictions
of the model to it.
```{r}
grid <- face %>%
  data_grid(d.comp) %>%
  add_predictions(fit)
head(grid)
```
Now we can plot the regression line and the original data just like any other plot.
```{r}
ggplot() +
  geom_point(data = face, mapping = aes(x = d.comp, y = diff.share)) +
  geom_line(data = grid, mapping = aes(x = d.comp, y = pred),
            colour = "red")
```
This method is more complicated than the `geom_abline` method for a bivariate regression, but will work for more complicated models, while the `geom_abline` method won't.


Note that `r ggdoc("geom_smooth")` can be used to add a regression line to a data-set.
```{r}
ggplot(data = face, mapping = aes(x = d.comp, y = diff.share)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
The argument `method = "lm"` specifies that the function `lm` is to be used to generate fitted values.
It is equivalent to running the regression `lm(y ~ x)` and plotting the regression line, where `y` and `x` are the aesthetics specified by the mappings.
The argument `se = FALSE` tells the function not to plot the confidence interval of the regression (discussed later).

### Regression towards the mean

### Merging Data Sets in R

See the `r R4DS` chapter [Relational data](http://r4ds.had.co.nz/relational-data.html).

```{r pres12,message=FALSE}
data("pres12", package = "qss")
```

To join both data frames
```{r results='hide'}
full_join(pres08, pres12, by = "state")
```
However, since there are duplicate names, `.x` and `.y` are appended.

**Challenge** What would happen if `by = "state"` was dropped? 

To avoid the duplicate names, or change them, you can rename before merging, 
```{r results='hide'}
full_join(select(pres08, state, Obama_08 = Obama, McCain_08 = McCain,
                 EV_08 = EV),
          select(pres12, state, Obama_12 = Obama, Romney_12 = Romney,
                 EV_12 = EV),
          by = "state")
```
or use the `suffix` argument to `full_join`
```{r}
pres <- full_join(pres08, pres12, by = "state", suffix = c("_08", "_12"))
head(pres)
```

**Challenge** Would you consider this data tidy? How would you make it tidy?

The **dplyr** equivalent functions for `r rdoc("base", "cbind")` is `r rdoc("dplyr", "bind_cols")`.
```{r}
pres <- pres %>%
  mutate(Obama2008_z = as.numeric(scale(Obama_08)),
         Obama2012_z = as.numeric(scale(Obama_12)))
```
Likewise, `r rdoc("dplyr", "bind_cols")` concatenates data frames by row.

We need to use the `as.numeric` function because `scale()` always returns a matrix.
Omitting `as.numeric()` would not produce an error in the code chunk above, since the columns of a data frame
can be matrices, but it would produce errors in some of the following code if it were omitted.

Scatter plot of states with vote shares in 2008 and 2012
```{r}
ggplot(pres, aes(x = Obama2008_z, y = Obama2012_z, label = state)) +
  geom_abline(colour = "white", size = 2) +
  geom_text() +
  coord_fixed() +
  scale_x_continuous("Obama's standardized vote share in 2008",
                     limits = c(-4, 4)) +
  scale_y_continuous("Obama's standardized vote share in 2012",
                     limits = c(-4, 4))
```

To calculate the bottom and top quartiles
```{r}
pres %>%
  filter(Obama2008_z < quantile(Obama2008_z, 0.25)) %>%
  summarise(improve = mean(Obama2012_z > Obama2008_z))

pres %>%
  filter(Obama2008_z < quantile(Obama2008_z, 0.75)) %>%
  summarise(improve = mean(Obama2012_z > Obama2008_z))
```

**Challenge:** Why is it important to standardize the vote shares?

### Model Fit

```{r message=FALSE}
data("florida", package = "qss")
fit2 <- lm(Buchanan00 ~ Perot96, data = florida)
fit2
```

Extract $R ^ 2$ from the results of `summary`,
```{r}
summary(fit2)$r.squared
```
Alternatively, can get the R squared value from the data frame `r rdoc("broom", "glance.lm", "glance")` returns:
```{r}
glance(fit2)
```

We can add predictions and residuals to the original data frame using the `r pkg_link("modelr")` functions `r rdoc("modelr", "add_residuals")` and `r rdoc("modelr", "add_predictions")`
```{r}
florida <-
  florida %>%
  add_predictions(fit2) %>%
  add_residuals(fit2)
glimpse(florida)
```
There are now two new columns in `florida`, `pred` with the fitted values (predictions), and `resid` with the residuals.

Use `fit2_augment` to create a residual plot:
```{r}
fit2_resid_plot <-
  ggplot(florida, aes(x = pred, y = resid)) +
  geom_ref_line(h = 0) +
  geom_point() +
  labs(x = "Fitted values", y = "residuals")
fit2_resid_plot
```
Note, we use the function `r rdoc("modelr", "geom_refline")` to add a reference line at 0.

Let's add some labels to points, who is that outlier?
```{r}
fit2_resid_plot +
  geom_label(aes(label = county))
```

The outlier county is "Palm Beach"
```{r}
arrange(florida) %>%
  arrange(desc(abs(resid))) %>%
  select(county, resid) %>%
  head()
```

Data without Palm Beach
```{r}
florida_pb <- filter(florida, county != "PalmBeach")
fit3 <- lm(Buchanan00 ~ Perot96, data = florida_pb)
fit3
```

$R^2$ or coefficient of determination
```{r}
glance(fit3)
```

```{r}
florida_pb %>%
  add_residuals(fit3) %>%
  add_predictions(fit3) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_ref_line(h = 0) +
  geom_point() +
  ylim(-750, 2500) +
  xlim(0, 1500) +
  labs(x = "Fitted values", y = "residuals")
```

Create predictions for both models using `r rdoc("modelr", "data_grid")` and `r rdoc("modelr", "gather_predictions")`:
```{r}
florida_grid <-
  florida %>%
  data_grid(Perot96) %>%
  gather_predictions(fit2, fit3) %>%
  mutate(model =
           fct_recode(model,
                      "Regression\n with Palm Beach" = "fit2",
                      "Regression\n without Palm Beach" = "fit3"))
```
Note this is an example of using non-syntactic column names in a tibble, as discussed in Chapter 10 of [R for data science](http://r4ds.had.co.nz/tibbles.html).

```{r}
ggplot() +
  geom_point(data = florida, mapping = aes(x = Perot96, y = Buchanan00)) +
  geom_line(data = florida_grid,
             mapping = aes(x = Perot96, y = pred,
                           colour = model)) +
  geom_label(data = filter(florida, county == "PalmBeach"),
             mapping = aes(x = Perot96, y = Buchanan00, label = county),
                           vjust = "top", hjust = "right") +
  geom_text(data = tibble(label = unique(florida_grid$model),
                           x = c(20000, 31000),
                           y = c(1000, 300)),
             mapping = aes(x = x, y = y, label = label, colour = label)) +
  labs(x = "Perot's Vote in 1996", y = "Buchanan's Votes in 1996") +
  theme(legend.position = "none")
```
See [Graphics for communication](http://r4ds.had.co.nz/graphics-for-communication.html#label) in *R for Data Science* on labels and annotations in plots.



## Regression and Causation


### Randomized Experiments

Load data
```{r message=FALSE}
data("women", package = "qss")
```

Proportion of female politicians in reserved GP vs. unreserved GP
```{r}
women %>%
  group_by(reserved) %>%
  summarise(prop_female = mean(female))
```

The diff in means estimator:
```{r}
# drinking water facilities

# irrigation facilities
mean(women$irrigation[women$reserved == 1]) -
    mean(women$irrigation[women$reserved == 0])
```

Mean values of `irrigation` and `water` in reserved and non-reserved districts.
```{r}
women %>%
  group_by(reserved) %>%
  summarise(irrigation = mean(irrigation),
            water = mean(water))
```

The difference between the two groups can be calculated with the function `r rdoc("base", "diff")`, which calculates the difference between subsequent observations.
This works as long as we are careful about which group is first or second.
```{r}
women %>%
  group_by(reserved) %>%
  summarise(irrigation = mean(irrigation),
            water = mean(water)) %>%
  summarise(diff_irrigation = diff(irrigation),
            diff_water = diff(water))
```

The other way uses **tidyr** `r rdoc("tidyr", "spread.lm", "spread")` and `r rdoc("tidyr", "gather.lm", "gather")`,
```{r}
women %>%
  group_by(reserved) %>%
  summarise(irrigation = mean(irrigation),
            water = mean(water)) %>%
  gather(variable, value, -reserved) %>%
  spread(reserved, value) %>%
  mutate(diff = `1` - `0`)
```
Now each row is an outcome variable of interest, and there are columns for the treatment (`1`) and control (`0`) groups, and the difference (`diff`).

```{r}
lm(water ~ reserved, data = women)
```
```{r}
lm(irrigation ~ reserved, data = women)
```

### Regression with multiple predictors

```{r message=FALSE}
data("social", package = "qss")
glimpse(social)
levels(social$messages)
fit <- lm(primary2006 ~ messages, data = social)
fit
```

Create indicator variables for each message:
```{r}
social <-
  social %>%
  mutate(Control = as.integer(messages == "Control"),
         Hawthorne = as.integer(messages == "Hawthorne"),
         Neighbors = as.integer(messages == "Neighbors"))
```
alternatively, create these using a for loop.
This is easier to understand and less prone to typos:
```{r}
for (i in unique(social$messages)) {
  social[[i]] <- as.integer(social[["messages"]] == i)
}
```
We created a variable for each level of `messages` even though we will exclude one of them.
```{r}
lm(primary2006 ~ Control + Hawthorne + Neighbors, data = social)
```

Create predictions for each unique value of `messages`
```{r}
unique_messages <-
  data_grid(social, messages) %>%
  add_predictions(fit)
unique_messages
```

Compare to the sample averages
```{r}
social %>%
  group_by(messages) %>%
  summarise(mean(primary2006))
```

Linear regression without intercept.
```{r}
fit.noint <- lm(primary2006 ~ -1 + messages, data = social)
fit.noint
```

Calculating the regression average effect is also easier if we make the control group the first level so all regression coefficients are comparisons to it.
Use `r rdoc("forcats", "fct_relevel")` to make "Control"
```{r}
fit.control <-
 mutate(social, messages = fct_relevel(messages, "Control")) %>%
 lm(primary2006 ~ messages, data = .)
fit.control
```

Difference in means
```{r}
social %>%
  group_by(messages) %>%
  summarise(primary2006 = mean(primary2006)) %>%
  mutate(Control = primary2006[messages == "Control"],
         diff = primary2006 - Control)
```

Adjusted R-squared is included in the output of `broom::glance()`
```{r}
glance(fit)
glance(fit)[["adj.r.squared"]]
```


### Heterogeneous Treatment Effects

Average treatment effect (ate) among those who voted in 2004 primary
```{r}
ate <-
  social %>%
  group_by(primary2004, messages) %>%
  summarise(primary2006 = mean(primary2006)) %>%
  spread(messages, primary2006) %>%
  mutate(ate_Neighbors = Neighbors - Control) %>%
  select(primary2004, Neighbors, Control, ate_Neighbors)
ate
```
Difference in ATE in 2004 voters and non-voters
```{r}
diff(ate$ate_Neighbors)
```


```{r}
social.neighbor <- social %>%
  filter( (messages == "Control") | (messages == "Neighbors"))

fit.int <- lm(primary2006 ~ primary2004 + messages + primary2004:messages,
              data = social.neighbor)
fit.int
```

```{r}
lm(primary2006 ~ primary2004 * messages, data = social.neighbor)
```

```{r}
social.neighbor <-
  social.neighbor %>%
  mutate(age = 2008 - yearofbirth)

summary(social.neighbor$age)

fit.age <- lm(primary2006 ~ age * messages, data = social.neighbor)
fit.age
```

Calculate average treatment effects
```{r}
ate.age <-
  crossing(age = seq(from = 25, to = 85, by = 20),
         messages = c("Neighbors", "Control")) %>%
  add_predictions(fit.age) %>%
  spread(messages, pred) %>%
  mutate(diff = Neighbors - Control)
ate.age
```

You can use `r rdoc("base", "poly")` function to calculate polynomials instead of adding each term, `age + I(age ^ 2)`.
Though note that the coefficients will be be different since by default `poly` calculates orthogonal polynomials instead of the natural (raw) polynomials.
However, you really shouldn't interpret the coefficients directly anyways, so this should matter.
```{r}
fit.age2 <- lm(primary2006 ~ poly(age, 2) * messages,
               data = social.neighbor)
fit.age2
```

Create a data frame of combinations of ages and messages using `r rdoc("modelr", "data_grid")`, which means that we only need to specify the variables, and not the specific values,
```{r}
y.hat <-
  data_grid(social.neighbor, age, messages) %>%
  add_predictions(fit.age2)
```

```{r}
ggplot(y.hat, aes(x = age, y = pred,
                  colour = str_c(messages, " condition"))) +
  geom_line() +
  labs(colour = "", y = "Predicted turnout rates") +
  theme(legend.position = "bottom")
```

```{r}
y.hat %>%
  spread(messages, pred) %>%
  mutate(ate = Neighbors - Control) %>%
  filter(age > 20, age < 90) %>%
  ggplot(aes(x = age, y = ate)) +
  geom_line() +
  labs(y = "Estimated average treatment effect",
       x = "Age") +
  ylim(0, 0.1)
```


### Regression Discontinuity Design

```{r message=FALSE}
data("MPs", package = "qss")

MPs_labour <- filter(MPs, party == "labour")
MPs_tory <- filter(MPs, party == "tory")

labour_fit1 <- lm(ln.net ~ margin,
                 data = filter(MPs_labour, margin < 0))
labour_fit2 <- lm(ln.net ~ margin, MPs_labour, margin > 0)

tory_fit1 <- lm(ln.net ~ margin,
                data = filter(MPs_tory, margin < 0))
tory_fit2 <- lm(ln.net ~ margin, data = filter(MPs_tory, margin > 0))
```

Use `r RDoc("modelr::data_grid")` to generate a grid for predictions.
```{r}
y1_labour <-
  filter(MPs_labour, margin < 0) %>%
  data_grid(margin) %>%
  add_predictions(labour_fit1)
y2_labour <-
  filter(MPs_labour, margin > 0) %>%
  data_grid(margin) %>%
  add_predictions(labour_fit2)

y1_tory <-
  filter(MPs_tory, margin < 0) %>%
  data_grid(margin) %>%
  add_predictions(tory_fit1)

y2_tory <-
  filter(MPs_tory, margin > 0) %>%
  data_grid(margin) %>%
  add_predictions(tory_fit2)
```

Tory politicians
```{r}
ggplot() +
  geom_ref_line(v = 0) +
  geom_point(data = MPs_tory,
             mapping = aes(x = margin, y = ln.net)) +
  geom_line(data = y1_tory,
            mapping = aes(x = margin, y = pred), colour = "red", size = 1.5) +
  geom_line(data = y2_tory,
            mapping = aes(x = margin, y = pred), colour = "red", size = 1.5) +
  labs(x = "margin of victory", y = "log net wealth at death",
       title = "labour")
```

We can actually produce this plot easily without running the regressions, by using `geom_smooth`:

```{r}
ggplot(mutate(MPs, winner = (margin > 0)),
       aes(x = margin, y = ln.net)) +
  geom_ref_line(v = 0) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE, mapping = aes(group = winner)) +
  facet_grid(party ~ .) +
  labs(x = "margin of victory", y = "log net wealth at death")
```

In the previous code, I didn't directly compute the the average net wealth at 0, so I'll need to do that here.
I'll use `r rdoc("modelr", "gather_predictions")` to add predictions for multiple models:
```{r}
spread_predictions(data_frame(margin = 0),
                   tory_fit1, tory_fit2) %>%
  mutate(rd_est = tory_fit2 - tory_fit1)
```


```{r}
tory_fit3 <- lm(margin.pre ~ margin, data = filter(MPs_tory, margin < 0))
tory_fit4 <- lm(margin.pre ~ margin, data = filter(MPs_tory, margin > 0))

(filter(tidy(tory_fit4), term == "(Intercept)")[["estimate"]] -
 filter(tidy(tory_fit3), term == "(Intercept)")[["estimate"]])
```
